# Hierarchical structure {#hierarchical}

## Making a hierarchical structure {#makestr}

We can use the `make_structure` function to creating nested and crossed hierarchical structures

```{r, echo=FALSE, message=FALSE,warning=FALSE}
plot_nested<-function(data_structure){
  par(mar=c(0,10,1,1))
  plot(NA, xlim=c(1,nrow(data_structure)), ylim=c(ncol(data_structure)+0.5,0.5), yaxt="n", ylab="",bty="n", xlab="", xaxt="n")
  axis(2,1:ncol(data_structure),colnames(data_structure), lty=0, las=2)
  for(i in 1:ncol(data_structure)){
    x<-data_structure[,i]
    line_coords <- sapply(1:max(x), function(j) range(which(x==j)))
    arrows(line_coords[1,],i,line_coords[2,],i, code=0, col=c(1,2))
    text(x=colMeans(line_coords),y=i-0.1,1:ncol(line_coords), cex=0.5)
  }
}
```

### Single Factor

Simplest structure - one grouping factor with multiple observations. Here we create a structure with 3 repeated observations of 10 individuals. The `structure` contains the name of the grouping factors and their sample sizes, and `repeat_obs` is the number of repeated observations.
```{r}
make_structure(structure="individual(10)", repeat_obs=3)
```


### Nested factors
If we want to have nested factors, so different hierarchical groups, where levels of one group only exist in one higher group then we can use the `/` symbol in the `structure` argument. For example, here we have 2 sexes, each with 10 individuals, with 3 repeated measurements each. 
```{r}
make_structure(structure="sex(2)/individual(10)", repeat_obs=3)
```

```{r,fig.width=6,fig.height=1.2, echo=FALSE}
dat_str <- make_structure(structure="sex(2)/individual(10)", repeat_obs=3)
plot_nested(dat_str)
```
Note that in the nesting, the sample size for the lower group now represents the number within each level of the higher, rather than the total sample size, so overall there is 20 individuals. 

We can nest as much as we want:
```{r}
make_structure(structure="species(2)/population(2)/individual(5)", repeat_obs=3)
```

```{r,fig.width=6,fig.height=1.8, echo=FALSE}
dat_str <- make_structure(structure="species(2)/population(2)/individual(5)", repeat_obs=3)
plot_nested(dat_str)
```


### Crossed factors
We can create completely crossed factors - every combination of levels exists - using the `+` symbol in the `structure` argument
```{r}
make_structure(structure="treatment(2) + individual(5)", repeat_obs=1)

```

We can combine crossed and nested structures:
```{r}
make_structure(structure="treatment(2) + sex(2)/individual(5)", repeat_obs=1)
```

We can also output the crossed and nested using `:`
```{r}
make_structure(structure="treatment(2) + individual(5) + treatment:individual", repeat_obs=1)
```


### Temporal structure
```{r}
make_structure(structure="year(2)/month(12)/day(30)", repeat_obs=1)
```

```{r}
make_structure(structure="year(2) + month(12) + day(30) + year:month:day", repeat_obs=1)
```

<!-- ```{r}
dat_str <- make_structure(structure="species5(2)/population(2)/individual(10)", repeat_obs=3)

plot_nested(dat_str)

structure="species(2)/population5(2)/individual(10)"
structure <- gsub("\\s","",structure)
gsub(".\\)(.)","\\1",structure)
strsplit((gsub("\\w+\\(\\d+\\)","",structure)),"")[[1]]



```
 -->





## Simulating data at different hierarchical levels
<!-- <div class="panel panel-success">
<div class="panel-heading">
**Biological example**
</div>
<div class="panel-body">
We have taken repeated measurements of adult body mass.  
</div>
</div> -->

<!-- I would replace “this is simulating “ to something more explicit, like “In essence, running mixed models requires to simulate datasets with repeated values for individuals. This can be done by simulating a single combined predictor at each level…” -->
In essence, this is simulating a single combined predictor at each level ($z$) with mean 0 and a given variance ($\sigma^2_z$). 

<div class="alert alert-info">

$$
y_{i,j} = \beta_0 + z_j + \epsilon_{i,j}
$$
$$
z \sim N(0,\sigma^2_z)
$$
$$
\epsilon \sim N(0,\sigma^2_\epsilon)
$$
</div>

At this point we need to specify some kind of hierarchical data structure. To make this we can use the `make_structure` function (which is still in development). The data structure is essentially a data.frame (or matrix), with all the grouping factors and their levels, as we would see in a typical dataset. To generate nested data with `make_structure` we use the `/` symbol and we specify the number of levels for each grouping factor with `(n)` - with nested data this represents the number of levels within each level of the higher grouping factor. We also specify how many times each combination is observed using `repeat_obs`. This function assumes everything is balanced.

```{r}
ds <- make_structure(structure = "sex(2)/individual(500)", repeat_obs=2)

head(ds)
```
In this case we have created a nested structure with 2000 observations - 2 observations for each of 1000 individuals, 500 for each of 2 sexes. We can then use this to simulate variation at different levels. Note that sample sizes are extracted from the data structure and so we do not need to specify `N` in the `simulate_population` function as above.

```{r}
squid_data <- simulate_population(
  data_structure = make_structure(structure = "sex(2)/individual(500)", repeat_obs=2),
  parameters = list(
    individual = list(
      vcov = 0.5
    ),
    residual = list(
      vcov = 0.5
    )
  )
)

data <- get_population_data(squid_data)
head(data)

library(lme4)
short_summary <- function(x) print(summary(x), correlation=FALSE, show.resids=FALSE, ranef.comp = c("Variance"))

short_summary(lmer(y ~ 1 + (1|individual), data))
```

Note, the names of the different levels in the parameter list need to exactly match those in the data structure. The order does not, however, have to be the same.

<!-- Dave: Also, in the code for the first example, how are the number of replicate observations within individuals determined? The observation level is not listed in the data_structure line.
 -->

We could also use an existing data structure. Here we can take the grouping factors and levels from an existing dataset and input them to `simulate_population`.
<!-- Anne: Short summary of what BTdata is (or link in the explanation to the BTdata page?) and why it is useful for this example (i.e. that you use the structure of the pedigree of this population), what you aim to test with it?
 -->

```{r}
library(MCMCglmm)
data(BTdata)
head(BTdata)

squid_data <- simulate_population(
  data_structure = BTdata[,c("dam","fosternest","sex")],
  parameters = list(
    dam = list(
      vcov = 0.2
    ),
    fosternest = list(
      vcov = 0.3
    ),
    residual = list(
      vcov = 0.5
    )
  )
)

data <- get_population_data(squid_data)
data

```


<!-- The importing a real dataset is very cool! But, I think there needs to be more explanation. How does this work when dam, fosternest, and residual lists are empty of variable names? Can you use the variances in the data instead of specifying them?
 -->
<!-- getME(mod,"theta")
library(numDeriv)
fm1Fun <- update(mod,devFunOnly=TRUE)
fm1_thpar <- getME(mod,"theta")
h <- hessian(fm1Fun, fm1_thpar)
sqrt(diag(solve(h)))

mySumm2 <- function(.) {
    c(beta=fixef(.),sigma=sigma(.), sig01=unlist(VarCorr(.)))
}
bootMer(mod,mySumm2,nsim = 100) -->

This setup also allows us to create variation at each level that is driven by specific predictors
<!-- 
Dave: At the bit just before Fixed Factors, what do you mean by “This setup”? It implies the data file import action in the previous section, which makes no sense. 
Anne: the sentence “This setup also allows us to create variation at each level that is driven by specific predictors” comes out of the blue, and I find the wording ‘this setup’ too unspecific.

-->

<div class="alert alert-info">

$$
y = \beta_0 + \boldsymbol Z_j \boldsymbol\beta_z + \boldsymbol X_i \boldsymbol\beta_x + \epsilon_{i,j}
$$
$$
\boldsymbol X \sim MVN(\mu_x,\Sigma_x)
$$
$$
\boldsymbol Z \sim MVN(\mu_z,\Sigma_z)
$$
$$
\epsilon \sim N(0,\sigma^2_\epsilon)
$$
</div>

```{r}
squid_data <- simulate_population(
  data_structure = make_structure(structure = "sex(2)/individual(500)", repeat_obs=2),
  parameters = list(
    individual = list(
      names = c("size","behaviour","physiology"),
      beta = c(0.1,0.3,0.2)
    ),
    observation = list(
      names=c("temperature","rainfall"),
      beta = c(0.2,-0.1)
    ),
    residual = list(
      vcov = 0.5
    )
  )
)

data <- get_population_data(squid_data)

short_summary(lmer(y ~ 1 + (1|individual), data))

short_summary(lmer(y ~ size + rainfall + (1|individual), data))

```




## Random slopes 

<div class="alert alert-info">

$$
y_{i,j} = \beta_0 + u_j + x_{i,j} \boldsymbol\beta_j + \epsilon_{i,j}
$$
$$
\begin{bmatrix} u_j \\\boldsymbol\beta_j \\ \end{bmatrix} \sim MVN(\begin{bmatrix} 0 \\ \bar{\beta} \\ \end{bmatrix},\Sigma_{u,\boldsymbol\beta})
$$
$$
\epsilon \sim N(0,\sigma^2_\epsilon)
$$
</div>
Can be seen as an interaction

We can specify random slopes by simulating a slopes variable (`ind_slope`) at the individual level ($\boldsymbol\beta_j$). Here we want to give the mean and variance of the slopes, rather than have the mean=0 and the slopes as deviations as in a LMM. 
```{r}

squid_data <- simulate_population(
  parameters = list(
    individual = list(
      names = c("ind_int","ind_slope"),
      mean =c(0,0.2), 
      beta=c(1,0),
      vcov =matrix(c(1,0.3,0.3,0.5),ncol=2,nrow=2)
    ),
    observation= list(
      names = c("environment"),
      vcov =c(1)
    ), 
    residual = list(
      names = c("residual"),
      vcov =c(0.5)
    ),
    interactions = list(
      names = c("ind_slope:environment"),
      beta =c(1)
    )
  ),
  data_structure=make_structure("individual(300)",repeat_obs=10)
)

data <- get_population_data(squid_data)

short_summary(lmer(y ~ environment + (1+environment|individual),data))
```



<!-- Can also implement it as a slope and deviations from the slope using `I()`, to fit in more with how it is modelled. 
 -->