# Genetic effects {#animal}
```{r, echo=FALSE, message=FALSE,warning=FALSE}
rm(list=ls())
# library(dplyr)
# library(flair)
# library(lme4)
# library(MCMCglmm)
# library(scales)
# library(knitr)
set.seed(25)
```
This vignette assumes that you are generally happy with how the `sim_population()` function works.

## Additive genetics effects {#va}
In order to simulate breeding values (additive genetic effects), we can provide the `simulate_population()` function with the relatedness structure in the population. The simplest way to do this is providing a pedigree using the the `pedigree` argument (a genetic relatedness matrix could also be given to the `cov_str` argument). The input to this argument needs to be a list, and the name of the pedigree in the list links it with the item in the parameter list.

**NOTE** the `simulate_population` function has very little error checking of pedigree structure at the moment 

When simulating breeding values, **all** individuals in pedigree need to be in the data_structure and *vice versa*. Having unsampled individuals (for example the base population) can be achieved in the sampling stage (not implemented yet). 

Lets start by importing a pedigree from the pedtricks package
```{r, eval=TRUE, message=FALSE,warning=FALSE}
library(squidSim)
library(nadiv)
library(gremlin)
library(pedtricks)

data(gryphons)

ped <- fix_ped(gryphons[,1:3])
head(ped)
names(ped)[1]<-"animal"

```

We can use this pedigree as a data_structure
```{r, eval=TRUE}
squid_data <- simulate_population(
  data_structure = ped[,1:3], 
 
  pedigree = list(animal=ped[,1:3]),
  
  parameters =list(
    animal = list(
      vcov = 0.2
    ),
    residual = list(
      vcov = 0.5
    )
  )
)

data <- get_population_data(squid_data)
head(data)
```

We can use the grelim pacakge to run a REML animal model
```{r, eval=TRUE, cache=TRUE}
Ainv <- makeAinv(ped)$Ainv

mod <- gremlin(y~1, random=~ animal,data=data,ginverse=list(animal=Ainv))
summary(mod)
```


We don't have to simulate a phenotype for everyone in the pedigree, so can include a subset of IDs in the data strcuture
```{r, eval=TRUE}
ds <- data.frame(animal =ped[sample(1:nrow(ped),1000,replace=FALSE),1])

squid_data <- simulate_population(
  data_structure = ds, 
 
  pedigree = list(animal=ped[,1:3]),
  
  parameters =list(
    animal = list(
      vcov = 0.2
    ),
    residual = list(
      vcov = 0.5
    )
  )
)

data <- get_population_data(squid_data)
head(data)
```

```{r, eval=TRUE, cache=TRUE}
Ainv <- makeAinv(ped)$Ainv

mod <- gremlin(y~1, random=~ animal,data=data,ginverse=list(animal=Ainv))
summary(mod)
```


We might want to simulate repeated measurements to allow estimation of permanent environment effects. The simplest way to do this is to create a duplicated column in the data structure of the individual IDs. Permanent environment effects that are not linked to the pedigree can then be simulated.

<!-- 
We might want to simulate repeated measurements to allow estimation of permanent environment effects. This is where being able to have something in the parameter list with a different name to the grouping factor is useful. In this way permanent environmental and additive genetic effects can be simulated in different parts of the parameter list, and linked to the same part of the data_structure.
 -->

:::: {.blackbox data-latex=""}
::: {.center data-latex=""}
**NOTICE!**
:::
The instructions given for simulating permanent environment effects using squidSim were incorrect in the vignette prior to version 0.2.0 (updated in September 2025). 
::::


```{r, eval=TRUE}

## make data structure with two observations per individual, with ID duplicated in two columns, animal and individual, and link the animal column in the data structure to the pedigree
ds <- data.frame(animal=rep(ped[,1], 2),individual=rep(ped[,1], 2))

squid_data <- simulate_population(
  data_structure = ds, 
  pedigree=list(animal=ped),
  parameters = list(
    animal = list(
      vcov = 0.2
    ),
    individual = list(
      vcov = 0.3
    ),
    residual = list(
      vcov = 0.5
    )
  )
)

data <- get_population_data(squid_data)
head(data)
```

```{r, eval=TRUE, cache=TRUE}
Ainv <- makeAinv(ped)$Ainv

mod_pe <- gremlin(y~1, random=~ animal + individual,data=data,ginverse=list(animal=Ainv))
summary(mod_pe)

```


## Multivariate genetic effects

We can simulate genetic effects affecting multiple phenotypes and the covariance between them, by specifying the number of response variables, and a covariance matrix, instead of only a variance.
```{r, eval=TRUE}

squid_data <- simulate_population(
  data_structure = ped,
  pedigree = list(animal = ped),
  n_response=2,
  parameters = list(
    animal = list(
      vcov = diag(2)

    ),
    residual = list(
      vcov = diag(2)
    )
  )
)

data <- get_population_data(squid_data)
head(data)
```


```{r a_multi, eval=TRUE, cache=TRUE}
library(MCMCglmm)
Ainv<-inverseA(ped)$Ainv
mod <- MCMCglmm(cbind(y1,y2)~1,
  random=~us(trait):animal, 
  rcov=~us(trait):units,
  data=data,
  ginverse=list(animal=Ainv),
  family=rep("gaussian",2),
  verbose=FALSE)
summary(mod)

```

<br>




## Sex specific genetic variance and inter-sexual genetic correlations

```{r,fig.width=10,fig.height=6}

ds <- data.frame(animal=gryphons[,"id"],sex=sample(c("Female","Male"),nrow(gryphons), replace=TRUE))

squid_data <- simulate_population(
  parameters = list(
    sex=list(
      fixed=TRUE,
      names=c("Female","Male"),
      beta=c(-0.5,0.5)
    ),
    animal= list(
      names = c("G_female","G_male"),
      vcov =matrix(c(0.1,-0.1,-0.1,0.4), nrow=2, ncol=2 ,byrow=TRUE)
      ),
    residual = list(
      names="residual",
      vcov = 0.1
    )
  ),
  data_structure = ds,
  pedigree = list(animal=ped),
  model = "y = Female + Male + I(Female)*G_female + I(Male)*G_male + residual"
)

data <- get_population_data(squid_data)
head(data)
par(mfrow=c(1,2))
boxplot(y~factor(sex),data)
plot(G_female~G_male,data)

```
<br>

## Indirect Genetic Effects {#IGE}
Indirect genetic effects are a bit more difficult to code. Lets take the example of maternal genetic effects. The maternal genetic effect that affects an individual's phenotype, is that of its mother, not itself. Here we can use `[]` to index the levels of the random effects within the formula. We need to be careful here as internally in `simulate_population()` the indexing of the factors in the data structure is done independently. We therefore need to generate a index for the mothers that links to the individual. We can do this using the `index_link` argument - in the code below we create a new factor to index with, called `dam_link`, that is the dam factor in our data structure, that has been indexed to match the animal factor.

Using this indexing trick, we can simulate the direct genetic and maternal genetic effects that an individual has (and the covariance between them), as well as generating an individual's phenotype from its own direct genetic effects, and its mother's maternal genetic effect.

```{r, eval=TRUE}
squid_data <- simulate_population(
  parameters=list(
    animal = list(
      names=c("direct","maternal"),
      vcov = matrix(c(1,0.3,0.3,0.5),2,2)
    ),
    residual = list(
      names="residual",
      vcov = 0.5
    )
  ),
  data_structure=ped,
  pedigree=list(animal=ped),
  index_link=list(dam_link="dam-animal"),
  model = "y = direct + maternal[dam_link] + residual"
)

data <- get_population_data(squid_data)

head(data)

```



## GxE

```{r, eval=TRUE}

squid_data <- simulate_population(
  parameters = list(
    animal = list(
      names = c("G_int","G_slope"),
      mean = c(0,0), 
      vcov = matrix(c(1,0.3,0.3,0.5),ncol=2,nrow=2,byrow=TRUE),
      beta = c(1,0)
    ),
    observation= list(
      names = c("environment"),
      vcov = c(0.2)
    ), 
    residual = list(
      names = c("residual"),
      vcov = c(0.5)
    ),
    interactions=list(
      names = "G_slope:environment",
      beta = 1
    )
  ),
  data_structure=rbind(ped,ped,ped,ped,ped),
  pedigree = list(animal=ped)
)

data <- get_population_data(squid_data)
library(lme4)
short_summary <- function(x) print(summary(x), correlation=FALSE, show.resids=FALSE, ranef.comp = c("Variance"))

short_summary(lmer(y ~ environment + (1+environment|animal),data))
```


## Dominance
Here we can make use of the dominance relatedness matrices that can be generated in the `nadiv` package

```{r, eval=TRUE, cache=TRUE}
data(warcolak)

pedD <- warcolak[,1:3]

Dmats <- makeD(pedD)
Dmat <- Dmats$D
```

We can input this into `simulate_population()` with the `cov_str` argument:
```{r, eval=TRUE}
ds <- data.frame(animal = pedD[,1], animalD = pedD[,1])

squid_data <- simulate_population(
  data_structure = ds, 
 
  pedigree = list(animal=pedD[,1:3]),
  cov_str = list(animalD=Dmat),

  parameters =list(
    animal = list(
      vcov = 0.2
    ),
    animalD = list(
      vcov = 0.2
    ),
    residual = list(
      vcov = 0.5
    )
  )
)

data <- get_population_data(squid_data)

```

We can then run a model to check the simulation, using {gremlin}, with inverse A and D matrices:
```{r, eval=TRUE, cache=TRUE}
Dinv <- Dmats$Dinv
Ainv_dom <- makeAinv(pedD)$Ainv

mod_dom <- gremlin(y~1, random=~ animal + animalD,data=data,ginverse=list(animal=Ainv_dom,animalD=Dinv))
summary(mod_dom)

```



## Inbreeding depression
Coming soon...


## Genetic Groups
Coming soon...

