
# Multi-response Models

We can simulate multiple phenotypes that covary at different levels. In the case of a simple random effects model, we can have a covariance matrix at each level, 

<!-- Anne: I would formulate this section “Multi-trait response”, as it is different from say multidimensional plasticity where a single trait can respond to several environmental variables. Also, perhaps add a short sentence of introduction here such as “We simulate here no longer one phenotype y varying across I individuals, but a number of phenotypes that (co)vary across individuals..”
 -->

<!-- 
$$
\begin{bmatrix} 
  y^{(1)} \\ 
  y^{(2)} 
\end{bmatrix}
=   \beta_0 + \boldsymbol Z_j + \boldsymbol\epsilon_{i,j}
$$ 
-->
<div class="alert alert-info">

$$
\boldsymbol Y_{i,j} =  \boldsymbol\beta_0 + \boldsymbol Z_j + \boldsymbol E_{i,j}
$$ 
$$
\boldsymbol Z \sim MVN(0,\Sigma_z)
$$
$$
\boldsymbol E \sim MVN(0,\Sigma_\epsilon)
$$
</div>

We can indicate that there are multiple phenotypes within the parameter list in two ways. First, we can use `n_response` in the parameter list, and specifying the covariance matrix (cov) at each level. In this way we can simulate covariance at each level. 

<!-- Anne: Should the blue background model not come after the first paragraph?
  We probably just need an intro paragraph
 -->

```{r}
squid_data <- simulate_population(
  data_structure=make_structure(structure = "individual(100)",repeat_obs=10),
  parameters=list(
    individual = list(
      vcov = matrix(c(1,0.5,0.5,1),nrow=2,ncol=2)
    ),
    residual = list(
      vcov = matrix(c(1,0.5,0.5,1),nrow = 2,ncol = 2)
    )
  ),
  N_response = 2
)

data <- get_population_data(squid_data)
head(data)

```

<!--Dave: I think in the Multi-response section there needs to be some explanation of the syntax, eg for lines like: cov = matrix(c(1,0.5,0.5,1),nrow=2,ncol=2). To what do each of those refer?

BOX
to code a matrix in R we use the `matrix` function. This takes a vector of values, and arranges then in a matrix, with dimensions specified with `nrow` and `ncol`. For big matrices this can be petty annoying. 
`Tri2M()` function from the package `MCMCglmm`

Tri2M(c(1,0.5,1,0.3,0.2,1), lower.tri = FALSE)



 -->



<!-- 
https://stackoverflow.com/questions/63007496/how-to-create-an-editable-matrix-in-shiny-app
make little shiny app that allows you to enter diagonal and 

 -->
Second, we can build up predictors at each level that drive this covariance. Here we make `beta` into a matrix ($B$), with predictors as rows, and responses as columns. 

<div class="alert alert-info">

$$
\boldsymbol Y_{i,j} = \beta_0 + \boldsymbol Z_j B_z + \boldsymbol X_i \boldsymbol B_x + \boldsymbol E_{i,j}
$$
$$
\boldsymbol X \sim MVN(\mu_x,\Sigma_x)
$$
$$
\boldsymbol Z \sim MVN(\mu_z,\Sigma_z)
$$
$$
\boldsymbol E \sim N(0,\Sigma_\epsilon)
$$
</div>

The formulation above (just random effects), can be simulated in a similar way with `beta` as an identity matrix (i.e. a predictor for each trait).
<!-- I got lost in the next section starting with: “The formulation above (just random effects), can be simulated in a similar way with beta as an identity matrix (i.e. a predictor for each trait).” might need more explanation.
 -->
```{r, eval=TRUE}

squid_data <- simulate_population(
  data_structure= make_structure(structure = "individual(100)",repeat_obs=10),
  parameters=list(
    individual = list(
      vcov =matrix(c(1,0.5,0.5,1),nrow=2,ncol=2),
      beta= diag(2)
    ),
    residual = list(
      vcov =matrix(c(1,0.5,0.5,1),nrow=2,ncol=2),
      beta= diag(2)
    )
  ),
  N_response=2
)

data <- get_population_data(squid_data)
head(data)

# library(MCMCglmm)
# mod <- MCMCglmm(cbind(y1,y2)~1,random=~us(trait):individual, rcov=~us(trait):units,data=data,family=rep("gaussian",2),verbose=FALSE)
# summary(mod)

```

Alternatively, you could also create multivariate phenotypes being affected by the same predictors. Here we have two phenotypes, affected by three predictors, and so we can create a 3x2 matrix of betas

```{r}
beta <- matrix(c(0.5,0.1,0.2,0.2,0.3,0.1),nrow=3,ncol=2)
beta
```

```{r, eval=TRUE}

squid_data <- simulate_population(
  data_structure= make_structure(structure = "individual(100)",repeat_obs=20),
  parameters= list(
    individual = list(
      vcov =matrix(c(1,0.5,0.5,1),nrow=2,ncol=2)
    ),
    observation = list(
      names = c("temperature", "rainfall", "wind"),
      beta= beta
    ),
    residual = list(
      vcov= matrix(c(1,0.5,0.5,1),nrow=2,ncol=2)
    )
  ),
  N_response=2
)

data <- get_population_data(squid_data)
head(data)

# library(MCMCglmm)
# mod <- MCMCglmm(cbind(y1,y2)~1,random=~us(trait):individual, rcov=~us(trait):units,data=data,family=rep("gaussian",2),verbose=FALSE)
# summary(mod)

```



```{r, eval=TRUE}

squid_data <- simulate_population(
  data_structure= make_structure(structure = "individual(100)",repeat_obs=20),
  
  parameters= list(
    individual = list(
      vcov =matrix(c(1,0.5,0.5,1),nrow=2,ncol=2)
    ),
    observation = list(
      names = c("temperature", "rainfall", "wind"),
      beta= beta
    ),
    interactions = list(
      names = c("temperature:rainfall"),
      beta=matrix(c(0.1,-0.3),ncol=2)
    ),
    residual = list(
      vcov= matrix(c(1,0.5,0.5,1),nrow=2,ncol=2)
    )
  ),
  N_response=2
)

data <- get_population_data(squid_data)
head(data)

# library(MCMCglmm)
# mod <- MCMCglmm(cbind(y1,y2)~1,random=~us(trait):individual, rcov=~us(trait):units,data=data,family=rep("gaussian",2),verbose=FALSE)
# summary(mod)

```

